---
title: "Basic Data Exploratory Analysis"
author: "Nguyen Trung Nam"
date: "`r Sys.Date()`"
output: html_document
number_sections: true
zoreto: True 
editor: 
  mode: source
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

LOAD LIBRARY & USER-DEFINED FUNCTIONS
```{r Library, echo=FALSE}
library(data.table)
library(tidyverse)
library(readr)
library(lubridate)
library(zoo)
library(forcats)
library(corrr)
library(moments)
```

# Introduction

Finding insightful information from data is also called the data exploration, which generally consists of 6 steps: 1. Formulating right questions 2. Data collection 3. Data munging 4. Basic exploratory data analysis 5. Advance exploratory data analysis 6. Model assessments.  

## Aim
In this exercise, the basic exploratory data analysis is carried out on a cleaned dataset. This focuses on understanding the dataset by using descriptive statistics and data visualization.

## Brief introduction on the basic exploratory data analysis

### Descriptive statistics
In this step several aspects of data are explored:  
- Data structure providing data dimension, data types  
- Calculate basic statistics

### Visualizing analysis

# Datasets used

```{r load data, echo=FALSE}
datset = read_csv("d:/Repository/Machine-Learning-with-R/basic_data_exploratory_analysis/aqidataset.csv")[,-1]
```
# Exploratoring data using descriptive statistics
```{r descriptive data analysis, echo=TRUE}
# explore data structure & data types
str(datset)
# summarize numeric variable: Overall Air Quality Index (AQI) value
summary(datset$`Overall AQI Value`)
# summarize numeric variable by cities
summary_df <- datset %>% group_by(city) %>%
              summarise(ave = mean(`Overall AQI Value`),
                        max = max(`Overall AQI Value`),
                        min = min(`Overall AQI Value`),
                        skew = skewness(`Overall AQI Value`),
                        q25 = quantile(`Overall AQI Value`, 0.25),   
                        median = median(`Overall AQI Value`),           
                        q75 = quantile(`Overall AQI Value`, 0.75),
                        sd  = sd(`Overall AQI Value`))
print(summary_df)
                        
                        
                        
```
# Exploratoring data using data visualization

# Reference
